---
title: "MachineLearningProjectAnalysis"
author: "Sudheer"
date: "October 12, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 

Load the required libraries in current session 

```{r library}
library(RCurl)
library(rpart) # Regressive Partitioning and Regression trees
library(rpart.plot) # Decision Tree plot
library(caret)
library(randomForest)
```

## Download the training and test data

Load training data and test data using url given in assignment 

```{r download_data}
#Download the data
#train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-train.csv"
#test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainingset <- read.csv("pml-training.csv", na.strings=c("", "NA"))
testingset <- read.csv("pml-testing.csv", na.strings=c("", "NA"))
```

## Check dimension of data

Check dimensions of data

```{r check_data}
# Check dimensions for number of variables and number of observations
#head(trainingset)
#head(testingset)
dim(trainingset)
dim(testingset)

```

## Clean the data

Remove redundant columns inside data

```{r clean_data}
# Delete columns with all missing values
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]

# Some variables are irrelevant to our current project: user_name, raw_timestamp_part_1, raw_timestamp_part_,2 cvtd_timestamp, new_window, and  num_window (columns 1 to 7). We can delete these variables.
trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```


Check dimensions after cleaning of data

```{r check_afterCleaning_data}
dim(trainingset)
dim(testingset)

```



## Divide the data into training and test data 

Create subtraining and sampling set 

```{r divide_data}
subsamples <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
dim(subTraining)
dim(subTesting)

```


Check class distribution

```{r check_class_distr}
plot(subTraining$classe, col="blue", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")

```


##Modellling using decision treee

```{r decision_tree}

model1 <- rpart(classe ~ ., data=subTraining, method="class")
# Predicting:
prediction1 <- predict(model1, subTesting, type = "class")
# Plot of the Decision Tree
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
confusionMatrix(prediction1, subTesting$classe)


```


##Modellling using random forest

```{r random_forest}

model2 <- randomForest(classe ~. , data=subTraining, method="class")
# Predicting:
prediction2 <- predict(model2, subTesting, type = "class")
# Test results on subTesting data set:
confusionMatrix(prediction2, subTesting$classe)

```

##Final Prediction 

```{r final_prediction}

predictfinal <- predict(model2, testingset, type="class")
predictfinal

```


